---
title: "Birthday"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Birthday}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```
## 1. Mathematical Derivation

The birthday problem asks: *what is the chance that at least two people in a group of $n$ share a birthday?*  

It’s easier to compute the complement: the chance all birthdays are different. There are $365 \times 364 \times \cdots \times (365-n+1)$ ways for this to happen, out of $365^n$ possible outcomes:

\[
P(\text{no match}) = \frac{365 \times 364 \times \cdots \times (365-n+1)}{365^n}.
\]

So the probability of a duplicate is:

\[
P(\text{duplicate}) = 1 - P(\text{no match}).
\]

---

## 2. Reformulation with `lfactorial` and `lchoose`

The numerator can be written as $\binom{365}{n} \times n!$, giving:

\[
P(\text{duplicate}) = 1 - \frac{\binom{365}{n} \times n!}{365^n}.
\]

Using logarithms, we express this in R as:

\[
P(\text{duplicate}) = 1 - \exp\big( \log \binom{365}{n} + \log(n!) - n \log(365) \big).
\]

---

## 3. The Function

This is implemented in R as:

```{r}
MATH4753F25ianh::birthday
```
